---
title: "individual2.Rmd"
author: "Seydou Yacouba"
date: "March 26, 2019"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
```

#load defense
```{r}
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)
```

#load offense
```{r}
Offensive <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive') 
str(Offensive)
```


#load biography
```{r}
biography <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
str(biography)
```


##The player names and opponent names are factors.
#All offensive and defensive statistics are numerical. 
#The Weight column in biogrpahy is also numeric. 
```{r}
defClean <- defense %>%
  mutate_at(c("Name", "Opponent_Opponent"), as.factor) %>% #1
  mutate_at(names(defense)[3:11], as.numeric) #2
str(defClean)
```
```{r}
offClean <- Offensive %>%
  mutate_at(c("Name", "Opponent_Opponent"), as.factor) %>% #1
  mutate_at(names(Offensive)[3:11], as.numeric) #1
str(offClean)
```
```{r}
bioClean <- biography %>%
  mutate_at(c("Name"), as.factor) %>% #1
  mutate_at(c("Weight"), as.numeric) #2
str(bioClean)
```


#Bonus
```{r}
bioClean <- bioClean %>%
  separate(Height, c("feet", "inch"), "-") %>%
  mutate_at(c("feet", "inch"), as.numeric) %>%
    mutate(new_height = feet + (inch / 12))
str(bioClean)
```
#Part two : tidying(1)
#1
#Reformat the defClean data frame into a tidy format using gather.
#The type of statistic (Tackles_Solo, Tackles_ASST, etc) is added as a 
#new key column named Statistic.
```{r}
defClean <- defClean %>% 
  gather(key = "Statistics", value = "Score", 3:11)
```

#2
#Create a histogram for each defensive statistic. 
#Use facet_wrap to draw all histograms in a single plot. 
#What defensive skills are rare?
```{r}
defClean %>%
  ggplot(aes(x = Score, fill=Statistics)) +
    geom_histogram() + 
    facet_wrap(~ Statistics)
```


#3
```{r}
defClean %>% 
  filter(Statistics == 'Tackles_Solo',
         Opponent_Opponent %in% c("West Virginia","Kansas State")) %>%
  spread(Opponent_Opponent, Score) %>%
  group_by(Name) %>%
  summarise(Total_kansas = sum(`Kansas State`),
            Total_west = sum(`West Virginia`)) %>%
  ggplot(aes(x=Total_west, y=Total_kansas)) + 
  geom_point() 
  
```


#part two: Tidying(2)
#1
```{r}

bioClean <- bioClean %>%
    separate(Hometown, c("City","State"), sep = ",")

head(bioClean)
```
#2
```{r}
bioClean %>%
  group_by(State) %>%
  count()
```




