---
title: "team"
output: html_document
---

##Clean data

```{r}
library("tidyverse")
library("readxl") # to import Excel file

defense <- read_excel("cyclonesFootball2018.xlsx", sheet="Defensive")
offense <- read_excel("cyclonesFootball2018.xlsx", sheet="Offensive")
biography <- read_excel("cyclonesFootball2018.xlsx", sheet="Biography")
```

# Player names and opponent names as factors
```{r}
defense$Name <- as.factor(defense$Name)
defense$Opponent_Opponent <- as.factor(defense$Opponent_Opponent)
offense$Name <- as.factor(offense$Name)
offense$Opponent_Opponent <- as.factor(offense$Opponent_Opponent)
biography$Name <- as.factor(biography$Name)
```

# All offensive and defensive statistics are numeric. Weight column in biography is also numeric.
```{r}
defense <- defense %>% mutate_at(c('Tackles_Solo','Tackles_ASST','Tackles_TFL','Tackles_Sack','Turnover_FF','Turnover_FR','Turnover_INT','Pass_QBH','Pass_PB'), as.numeric)
```

```{r}
offense <- offense %>% mutate_at(c('Receiving_REC','Receiving_YDS','Receiving_TD','Rushing_ATT','Rushing_YDS','Rushing_TD','Passing_CMP-ATT','Passing_YDS'), as.numeric)
```
```{r}
biography$Weight <- as.numeric(biography$Weight)
```


# Change Height column in biography into numeric (using a unit that works well)
```{r}
biography$Height <- sapply(strsplit(as.character(biography$Height), '-'), function(x){12*as.numeric(x[1]) + as.numeric(x[2])})
```
# The measurement is now in total inches

```{r}
defClean <- defense
offClean <- offense
bioClean <- biography
data <- defClean
```

##Part Two: Tidying
```{r}
defClean <- gather(defClean, key='Statistic', value='Amount', Tackles_Solo:Pass_PB, na.rm=TRUE)

p <- ggplot(defClean, aes(x=Amount)) + geom_histogram(bins=10) + facet_wrap(~defClean$Statistic)
p <- p + xlab("Value of Statistic") + ylab('Occurences') + ggtitle('Amount of Occurences per Defensive Statistic')
p

data <- data %>% select(Name, Opponent_Opponent, Tackles_Solo) %>% spread(Opponent_Opponent, Tackles_Solo, fill=0)
data <- data %>% select(Name, 'West Virginia', 'Kansas State')

p <- ggplot(data, aes(x=`West Virginia`, y=`Kansas State`)) + geom_point() + geom_smooth(method=lm, se=FALSE)
p <- p + xlab("West Virginia") + ylab("Kansas State") + ggtitle("Solo Tackles of WVU vs. KSU")
p

bioClean <- separate(bioClean, Hometown, c('City', 'State'), ', ')
head(bioClean, 3)

p <- ggplot(bioClean, aes(x=State, color=State)) + geom_bar() + theme(axis.text.x = element_text(angle = 90))
p <- p + ylab('Number of Players') + ggtitle('Number of ISU Football Players by State')
p
```

##Part Three: Joining Data Frames
```{r}
combo <- merge(bioClean, offClean, by='Name')
```

# Find where offensive players come from by joining offensive data and biography data appropriately
```{r}
p <- ggplot(combo, aes(x=State, color=State)) + geom_bar() + theme(axis.text.x = element_text(angle = 90))
p <- p + ylab('Number of Players') + ggtitle('Number of Offensive Players by State')
p
```


#2

```{r}
ggplot(combo, aes(x=Weight, y=Receiving_YDS)) +
  geom_jitter() 
  
```

# There is no relationship between the weight of a player and the receiving yards, we can #that some players that are less than 180 have less receiving yard while some witght a weight #over 220 have a high receiving yard. 


#3
# what's the number of Receiving_YDS in each game?
```{r}
oSumm <- offense %>%
  group_by(Name) %>% 
  summarize(receiving = sum(Receiving_YDS)) 
head(oSumm)
```


